Meta-Analysis of the `Ironic' Effects of Intergroup Contact

### 2022-02-17 manuscript/manuscript.tex
revision

\documentclass[twocolumn, 11pt, letterpaper]{article}

% Font
\usepackage[utf8]{inputenc}
\usepackage{MinionPro} 
\input{glyphtounicode}
\pdfgentounicode=1 
\usepackage{microtype}

% Format
\usepackage[letterpaper, margin = 1in]{geometry}
\setlength{\columnsep}{18pt}
\setcounter{secnumdepth}{0}
\usepackage[compact]{titlesec}  
\titleformat*{\section}{\centering\normalfont\Large\bfseries}
\usepackage{authblk}
\renewcommand\Affilfont{\small}
\newcommand{\refsection}{\section{References}}
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

% References
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{CSLReferences}%
  {}%
  {\par}

% Frontmatter
\title{ Meta-Analysis of the `Ironic' Effects of Intergroup
Contact\thanks{} }
\author[1]{Nils K. Reimer}
\author[2]{Nikhil K. Sengupta}
\affil[1]{University of Southern California, Los Angeles, United States}
\affil[2]{University of Kent, Canterbury, United Kingdom}
\date{June 15, 2021}

@@

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
      \noindent Growing evidence suggests that intergroup contact,
psychology's most-researched paradigm for reducing prejudice, has the
`ironic' effect of reducing support for social change in disadvantaged
groups. We conducted a preregistered meta-analytic test of this effect
across 98 studies with 140 samples of 213,085 disadvantaged-group
members. As predicted, intergroup contact was, on average, associated
with less perceived injustice (\(r = -.07\)), collective action
(\(r = -.06\)), and support for reparative policies (\(r = -.07\)).
However, these associations were small, variable, and consistent with
alternative explanations. Across outcomes, 25--36\% of studies found
\emph{positive} associations with intergroup contact. Moderator analyses
explained about a third of the between-sample variance, showing that, on
average, associations were negative only in studies of adults that
measured intergroup contact directly and were strongest in studies that
examined short-term migration or (post-)colonial intergroup relations.
We also found evidence for an alternative explanation for the apparent
`ironic' effects of intergroup contact as, after controlling for the
positive association of negative contact with support for social change,
positive contact was no longer associated with any of the outcomes. We
close by discussing strengths and limitations of the available evidence
and by highlighting open questions about the relationship between
intergroup contact and support for social change in disadvantaged
groups.
    \end{abstract}
  \end{@twocolumnfalse}
  % \renewcommand{\abstractname}{Statement of Relevance}
  % \begin{@twocolumnfalse}
  %   \maketitle
  %   \begin{abstract}
  %   \noindent 
  %   \end{abstract}
  % \end{@twocolumnfalse}
]

{
  \renewcommand{\thefootnote}%
    {\fnsymbol{footnote}}
  \footnotetext[1]{Manuscript under review.}
}

\noindent It is ironic that psychology's most-researched paradigm for
reducing prejudice---intergroup contact---should hinder, rather than
help, social change. Hundreds of studies have confirmed that intergroup
contact is associated with less negative feelings toward outgroup
members (for a meta-analysis, see Pettigrew \& Tropp, 2006). Most of
these studies have focused on advantaged-group members' contact
experiences with disadvantaged-group members (L. R. Tropp \& Pettigrew,
2005), guided by the idea that reducing advantaged-group members'
prejudice will prevent interpersonal discrimination. Reducing prejudice,
however, is rarely enough to overcome institutional and structural
discrimination. Instead, as the 2020 Black Lives Matter protests have
shown, social change often requires political mobilization by the
disadvantaged. Critics have argued that, by fostering harmony and
reducing conflict, intergroup contact might distract disadvantaged-group
members from the injustice they face and, thereby, undermine their
motivation for social change (Dixon et al., 2005; Reicher, 2007; Wright,
2003). By reducing disadvantaged-group members' opposition to injustice,
intergroup contact might thus have the `ironic' (Saguy et al., 2009),
`sedative' (Çakal et al., 2011), or `paradoxical' (Dixon et al., 2010)
effect of hindering social change.

@@

Collective action is any action directed at improving the conditions of
the disadvantaged ingroup (Wright et al., 1990). Collective action is
crucial to achieving social change because advantaged-group members
rarely give up their advantages without sustained pressure from
disadvantaged-group members (Blumer, 1958; for a review, see Dixon et
al., 2012). Researchers have argued that intergroup contact could
discourage disadvantaged-group members from engaging in collective
action by diminishing their perceived discrimination (Linda R. Tropp et
al., 2012), by blurring boundaries between the disadvantaged ingroup and
the advantaged outgroup (Saguy et al., 2009), by reducing negative
attitudes toward the advantaged outgroup (Wright \& Lubensky, 2009), and
by quelling their anger (Hayward et al., 2018; Tausch et al., 2015). To
the extent that it discourages collective action, intergroup contact
reduces support for social change.

@@

While narrative reviews of the growing literature on this topic have
been published (Hässler et al., 2021; MacInnis \& Hodson, 2019; McKeown
\& Dixon, 2017), we present the first systematic review and
meta-analysis of the relationship between intergroup contact and support
for social change in disadvantaged groups. To test the hypothesized
`ironic' effects of intergroup contact, we evaluate the evidence that
contact with advantaged-group members reduces disadvantaged-group
members' perceived injustice, collective action, and support for
reparative policies.

Our first objective was to systematically review the available evidence
and to evaluate its strengths and limitations. Our second objective was
to synthesize the available evidence to establish the direction and
magnitude of the relationships between intergroup contact and the three
outcomes. Doing so allowed us to determine whether the available
evidence supports a negative or positive (see Pettigrew et al., 2011)
relationship between contact and support for social change and to
estimate the strength of this relationship. Our third objective was to
estimate and explain the variability of this relationship across study
settings, designs, and other potential moderators. For example, we
examined whether any effects were confined to Western, Educated,
Industrialized, Rich, Democratic (WEIRD) countries, the cultural context
of most psychological research (Henrich et al., 2010). Our fourth
objective was to assess how robust the available evidence was to
publication bias. Our final objective was to explore the available
evidence for negative contact and ingroup contact as alternative
explanations of the apparent `ironic' effects of intergroup contact.

@@

We prepared a protocol in accordance with the PRISMA-P statement (Moher
et al., 2015) and preregistered it on the Open Science
Framework.\footnote{\url{https://osf.io/ryuev/?view_only=8938ac6178914da98bdda1da4a6e27f0}}
Preregistration precludes undisclosed flexibility in study selection,
outcome selection, and other decisions that influence effect size
estimates---and thus prevents confirmation bias from determining the
conclusions of a meta-analysis (Lakens et al., 2016). We report
deviations from our protocol in the Supplemental Online Material (SOM).
We make data, analysis scripts, and the fully reproducible manuscript
available online.\footnote{\url{https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a}}

@@

As preregistered, we searched titles, abstract, and keywords for
relevant terms in four electronic databases. We searched for relevant
articles in the \emph{Scopus} and \emph{PsycINFO} databases. We searched
for gray literature in the \emph{ProQuest Dissertations and Theses}
database. We used similar non-exclusive search terms for all databases
(see SOM). We searched databases on April 1, 2019 and again on April 1,
2020. We exported records and relevant metadata from each database. We
removed duplicates using the \emph{revtools} package (Westgate, 2019).

@@

\hypertarget{outcomes-selection}{%
\subsubsection{Outcomes Selection}\label{outcomes-selection}}

@@

\textbf{\emph{Study Setting.}} For each sample, we recorded in what
country (or countries) the data was collected (\(\kappa = .97\)) and
what the disadvantaged ingroup and the advantaged outgroup were. We
categorized each sample's setting according to whether the source of the
groups' relative inequality was long-term migration (e.g., Asian
Americans), short-term migration (e.g., international students), slavery
(e.g., Black Americans), colonization (e.g., Māori, Black South
Africans), religion, caste, sexuality, or another distinction
(\(\kappa = .85\)).

\textbf{\emph{Study Design.}} We categorized each study as either
observational and cross-sectional, observational and longitudinal,
quasi-experimental, experimental, an intervention, or other
(\(\kappa = .85\)). We categorized each sample as either a student
convenience sample, a non-student convenience sample, a probability or
representative sample, or another kind of sample (\(\kappa = .73\)). We
categorized the age group(s) in each sample as children (\(\leq 12\)
years), adolescents (13--18 years), or adults (\(\geq 18\) years;
\(\kappa = .64\)).

@@

\textbf{\emph{Predictor Variables.}} For each measure of intergroup
contact, we coded whether it assessed contact with the specific
advantaged outgroup of interest directly---or whether it assessed
contact indirectly by, for example, asking what proportion of someone's
friends were not from the participants' ingroup or what proportion of
residents in someone's neighbourhood were from the relevant outgroup
(\(\kappa = .82\)).

@@

We estimated effect sizes using Bayesian random-effects meta-analysis
models in \emph{RStan} (Stan Development Team, 2020) which modeled the
\emph{z}-transformed correlation coefficients with a normal likelihood
function:
\begin{align*} z_{ij} &\sim \text{Normal}(\theta_{ij}, \sigma_{ij}) \\ \theta_{ij} &= \begin{cases} \mu + \beta_j\tau_J & \text{if } I_j = 1 \\ \mu + \beta_j\tau_J + \beta_i\tau_I & \text{if } I_j > 1 \end{cases} \\ \end{align*}
where \(z_{ij}\) is the observed effect size in sample \(i\) of study
\(j\), \(\sigma_{ij}\) is the sample standard error, and \(\theta_{ij}\)
is the estimated effect size. We estimated \(z_{ij}\) as a function of
the mean effect size \(\mu\) and of two varying (random) intercepts,
\(\beta_i\) and \(\beta_j\), with the corresponding standard deviations,
\(\tau_I\) and \(\tau_J\). We used the non-centered parameterization to
model the random effects. For studies that contained only one sample
(\(I_j = 1\)), we only included \(\beta_j\tau_J\), the study-wise
deviation from the mean effect size \(\mu\). For studies that contained
more than one sample (\(I_j > 1\)), we also estimated \(\beta_i\tau_I\),
the sample-wise deviation from the study-specific effect size.

@@

Bayesian inference involves choosing a likelihood function and prior
distributions. A likelihood function links the observed data to one or
more model parameters and states how likely the observed data are given
different values of said model parameters. Prior distributions state how
plausible different values of said model parameters are before
considering the observed data. Bayesian inference applies Bayes' theorem
to update prior distributions in light of the observed data to produce
posterior distributions. In contrast to \emph{p}-values and confidence
intervals, the resulting posterior distributions have a straightforward
interpretation as stating how plausible different values of the model
parameters are given the observed data. We report point estimates, based
on the median of posterior samples, and uncertainty intervals, based on
the quantiles of posterior samples, that enclose the 95\% most plausible
estimates. In addition, we report the posterior probability, based on
the proportion of posterior samples below zero, that the mean effect
size is negative.

@@

@@

As preregistered, we conducted two kinds of robustness checks. First, we
assessed to what extent our findings were sensitive to choosing
narrower, \(\mu \sim \text{Normal}(0, 0.1)\), or wider,
\(\mu \sim \text{Normal}(0, 1)\), prior distributions. Choosing narrower
or wider prior distribution did not affect mean effect size estimates
for perceived injustice (\(\Delta r = -.00, [-.05, .04]\) and
\(\Delta r = .00, [-.05, .05]\)), collective action
(\(\Delta r = -.01, [-.11, .09]\) and \(\Delta r = .00, [-.11, .10]\)),
and policy support (\(\Delta r = -.01, [-.10, .08]\) and
\(\Delta r = .00, [-.09, .10]\)). Second, we assessed to what extent our
findings were sensitive to including or excluding influential studies by
repeating the preregistered analyses \(J\) times while leaving out one
of \(J\) studies each time and by calculating the mean absolute
difference (\emph{MAD}) for the estimated mean effect size across
left-out studies. For perceived injustice
(\(\textit{MAD} = .02, [.01, .04]\)), collective action
(\(\textit{MAD} = .04, [.02, .09]\)), and policy support
(\(\textit{MAD} = .03, [.02, .08]\)), the \emph{MAD} was small. Leaving
out the most influential study, for example, did not change estimates of
the mean effect size for the three outcomes
(\(\Delta r = -.00, [-.05, .04]\); \(\Delta r = .02, [-.08, .12]\);
\(\Delta r = -.02, [-.11, .07]\)). These analyses showed that our
findings were robust to choosing different prior distributions and to
excluding influential studies.

@@

\begin{figure*}
\centering
\caption{Estimated effect size as a function of cultural distance from the United States, with point estimates and uncertainty intervals for each country}
\includegraphics[scale=1]{../figures/figure-5}
\caption*{\textit{Note.} We had data from too few countries to reach firm conclusions about the direction of these associations for perceived injustice ($\beta = -0.38, [-1.45, 0.72]$; $\Pr (\beta < 0) = 75.7\%$), collective action ($\beta = -0.09, [-1.60, 1.37]$; $\Pr (\beta < 0) = 55.1\%$), and policy support ($\beta = -0.92, [-2.13, 0.55]$; $\Pr (\beta < 0) = 90.3\%$).}
\label{fig:f5}
\end{figure*}

@@

First, we used meta-regression models to examine categorical moderator
variables which we expected to explain heterogeneity in effect sizes
across samples. As Figure 4 shows, we found that the setting of the
study (\(R^2 = 20\%\)), participants' age group (\(R^2 = 11\%\)), and
whether intergroup contact was measured directly or indirectly
(\(R^2 = 10\%\)) explained the most variance across effect sizes.

@@

Third, we used meta-regression trees to discover interactions between
moderator variables that best explained heterogeneity in effect sizes
(Li et al., 2017, 2020). Figure 6 shows the resulting meta-regression
model, which explained more variance across samples than any individual
moderator (\(R^2 = 31\%\)). We found that intergroup contact was
associated with less perceived injustice only in studies that focused on
adults and that measured intergroup contact directly. Among these
studies, this association was stronger in settings in which the groups'
relative inequality stemmed from short-term migration or colonization
(\(r = -.18, [-.23, -.13]\)) than in other settings
(\(r = -.09, [-.13, -.06]\)).

@@

Developing methods to detect and correct for publication bias and other
meta-biases is an active area of research, with no single method
outperforming all others (Carter et al., 2019). Following Carter et
al.'s recommendations, we compared results of several methods to adjust
meta-analytic estimates for publication bias: the PET-PEESE estimator
(Stanley \& Doucouliagos, 2014); the three-parameter selection model
{[}3PSM; Vevea \& Hedges (1995){]}; and the \emph{p}-uniform* estimator
(van Aert \& Assen, 2018). In addition, we ran a subgroup analysis that
included only studies that were not published or not conducted with the
intention to examine the effects of intergroup contact on any of the
outcomes---and that, therefore, can be assumed to have not been affected
by publication bias. As these methods do not use Bayesian statistics, we
compared adjusted estimates to unadjusted estimates from a
random-effects meta-analysis estimated with restricted maximum
likelihood estimation using the \emph{metafor} package (Viechtbauer,
2010).

@@

\hypertarget{alternative-predictors}{%
\subsection{Alternative Predictors}\label{alternative-predictors}}

\begin{figure*}[t!]
\centering
\caption{Posterior distributions from the random-effects meta-analysis models with alternative predictor variables}
\includegraphics[scale=1]{../figures/figure-8}
\label{fig:f8}
\end{figure*}

>>>>

\documentclass[12pt, letterpaper]{article}

% Font
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{microtype}

% Format
\usepackage[letterpaper, margin = 1in]{geometry}
\setcounter{secnumdepth}{0}
\setlength{\parindent}{0.5in}
\usepackage[compact]{titlesec}
\titleformat*{\section}{\centering\normalfont\normalsize\bfseries}
\titleformat*{\subsection}{\raggedright\normalfont\normalsize\bfseries}
\titleformat*{\subsubsection}{\raggedright\normalfont\normalsize\bfseries\itshape}
\newcommand{\refsection}{\newpage \section{References}}
\usepackage{indentfirst}
\usepackage[hang,flushmargin]{footmisc}
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Header
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{14.5pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{`IRONIC' EFFECTS META-ANALYSIS}
\fancyhead[R]{\thepage}

@@

% References
\newenvironment{CSLReferences}[2]{}{}

% Line spacing
\usepackage[doublespacing]{setspace}

% Frontmatter
\title{ Meta-Analysis of the `Ironic' Effects of Intergroup Contact }

@@

\section{Abstract}

\noindent Growing evidence suggests that intergroup contact,
psychology's most-researched paradigm for reducing prejudice, has the
`ironic' effect of reducing support for social change in disadvantaged
groups. We conducted a preregistered meta-analytic test of this effect
across 98 studies with 140 samples of 213,085 disadvantaged-group
members. As predicted, intergroup contact was, on average, associated
with less perceived injustice (\(r = -.07\)), collective action
(\(r = -.06\)), and support for reparative policies (\(r = -.07\)).
However, these associations were small, variable, and consistent with
alternative explanations. Across outcomes, 25--36\% of studies found
\emph{positive} associations with intergroup contact. Moderator analyses
explained about a third of the between-sample variance, showing that, at
least for perceived injustice, associations with intergroup contact were
most consistently negative in studies that measured direct,
qualitatively positive contact among adults. We also found evidence for
an alternative explanation for the apparent `ironic' effects of
intergroup contact as, after controlling for the positive association of
negative contact with support for social change, positive contact was no
longer associated with any of the outcomes. We close by discussing
strengths and limitations of the available evidence and by highlighting
open questions about the relationship between intergroup contact and
support for social change in disadvantaged groups.

\textit{Keywords:} intergroup contact, meta-analysis, perceived
discrimination, collective action, policy support

\newpage

\section{Meta-Analysis of the `Ironic' Effects of Intergroup Contact}

It is ironic that psychology's most-researched paradigm for reducing
prejudice---intergroup contact---should hinder, rather than help, social
change. Hundreds of studies have confirmed that intergroup contact is
associated with less negative feelings toward outgroup members (for a
meta-analysis, see Pettigrew \& Tropp, 2006). Most of these studies have
focused on advantaged-group members' contact experiences with
disadvantaged-group members (Tropp \& Pettigrew, 2005), guided by the
idea that reducing advantaged-group members' prejudice will prevent
interpersonal discrimination. Reducing prejudice, however, is rarely
enough to overcome institutional and structural discrimination. Instead,
as the 2020 Black Lives Matter protests have shown, social change often
requires political mobilization by the disadvantaged. Critics have
argued that, by fostering harmony and reducing conflict, intergroup
contact might distract disadvantaged-group members from the injustice
they face and, thereby, undermine their motivation for social change
(Dixon et al., 2005; Reicher, 2007; Wright, 2003). By reducing
disadvantaged-group members' opposition to injustice, intergroup contact
might thus have the `ironic' (Saguy et al., 2009), `sedative' (Çakal et
al., 2011), or `paradoxical' (Dixon et al., 2010) effect of hindering
social change.

@@

Collective action is any action directed at improving the conditions of
the disadvantaged ingroup (Wright et al., 1990). Collective action is
crucial to achieving social change because advantaged-group members
rarely give up their advantages without sustained pressure from
disadvantaged-group members (Blumer, 1958; for a review, see Dixon et
al., 2012). Researchers have argued that intergroup contact could
discourage disadvantaged-group members from engaging in collective
action by diminishing their perceived discrimination (Tropp et al.,
2012), by reducing the perceived threat from the advantaged outgroup
(Çakal et al., 2016), by blurring boundaries between the disadvantaged
ingroup and the advantaged outgroup (Saguy et al., 2009), by reducing
negative attitudes toward the advantaged outgroup (Wright \& Lubensky,
2009), and by quelling their anger (Hayward et al., 2018; Tausch et al.,
2015). To the extent that it discourages collective action, intergroup
contact reduces support for social change.

@@

As a result of two recent developments, there is now an emerging
consensus that, in line with the former argument, intergroup contact has
the `ironic' effect of undermining support for social change in
disadvantaged groups. First, a large-scale preregistered multi-country
study provided the most comprehensive evidence to date that contact with
advantaged groups correlates with less support for social change in
disadvantaged groups (Hässler et al., 2020). Second, several narrative
reviews of the growing evidence for the `ironic' effects of intergroup
contact have been published (Hässler, Uluğ, et al., 2021; MacInnis \&
Hodson, 2019; McKeown \& Dixon, 2017). Narrative reviews do not,
however, include all relevant studies and do not synthesize the,
sometimes conflicting, findings of different studies. We present the
first systematic review and meta-analysis of the relationship between
intergroup contact and support for social change in disadvantaged
groups. To test the hypothesized `ironic' effects of intergroup contact,
we evaluate the evidence that contact with advantaged-group members
reduces disadvantaged-group members' perceived injustice, collective
action, and support for reparative policies.

Our first objective was to systematically review the available evidence
and to evaluate its strengths and limitations. Our second objective was
to synthesize the available evidence to establish the direction and
magnitude of the relationships between intergroup contact and the three
outcomes. Doing so allowed us to determine whether the available
evidence supports a negative or positive direct relationship between
contact and support for social change and to estimate the strength of
this relationship. Our third objective was to estimate and explain the
variability of this relationship across study settings, designs, and
other potential moderators. For example, we examined whether any effects
were confined to Western, Educated, Industrialized, Rich, Democratic
(WEIRD) countries, the cultural context of most psychological research
(Henrich et al., 2010). Our fourth objective was to assess how robust
the available evidence was to publication bias. Our final objective was
to explore the available evidence for negative contact and ingroup
contact as alternative explanations of the apparent `ironic' effects of
intergroup contact.

In pursuing these objectives, we focus on the direct relationships of
intergroup contact with perceived injustice, collective action, and
support for reparative policies in disadvantaged groups. We do not
consider indirect and interaction effects or the effects of imagined,
extended, or parasocial contact. Further, while the argument for the
`ironic' effects of intergroup contact is part of a broader critique of
the prejudice reduction model of social change (for a review, see Dixon
et al., 2012), our meta-analysis tests the narrower hypothesis whether
intergroup contact reduces support for social change in disadvantaged
groups.

@@

We prepared a protocol in accordance with the PRISMA-P statement (Moher
et al., 2015) and preregistered it on the Open Science
Framework.\footnote{\url{https://osf.io/ryuev}} Preregistration
precludes undisclosed flexibility in study selection, outcome selection,
and other decisions that influence effect size estimates---and thus
prevents confirmation bias from determining the conclusions of a
meta-analysis (Lakens et al., 2016). We report deviations from our
preregistered protocol at the end of the Method section. We make data,
analysis scripts, and the fully reproducible manuscript available
online.\footnote{\url{https://osf.io/w5tqv}}

@@

As preregistered, we searched titles, abstract, and keywords for
relevant terms in four electronic databases. We searched for relevant
articles in the \emph{Scopus} and \emph{PsycINFO} databases. We searched
for gray literature in the \emph{ProQuest Dissertations and Theses}
database. We used similar non-exclusive search terms for all databases
(see SOM). We searched databases on April 1, 2019 and again on April 1,
2020. We exported records and relevant metadata from each database. We
removed duplicates using the \emph{revtools} R package (Westgate, 2019).

@@

\hypertarget{outcome-selection}{%
\subsubsection{Outcome Selection}\label{outcome-selection}}

@@

\textbf{\emph{Study Setting.}} For each sample, we recorded in what
country (or countries) the data was collected (\(\kappa = .97\)) and
what the disadvantaged ingroup and the advantaged outgroup were. We
categorized each sample's setting according to whether the source of the
groups' relative inequality was long-term migration (e.g., Asian
Americans), short-term migration (e.g., international students), slavery
(e.g., Black Americans), colonization (e.g., Māori, Black South
Africans), religion, caste, sexual/gender identity, or another
distinction (\(\kappa = .85\); see SOM for details).

\textbf{\emph{Study Design.}} We categorized each study as either
observational and cross-sectional, observational and longitudinal,
quasi-experimental, experimental, an intervention, or other
(\(\kappa = .85\)).\footnote{While the term `observational study' is
  sometimes used to refer to the naturalistic observation of behavior,
  we use it in its broader sense, that is, to refer to studies that do
  not involve random assignment into experimental conditions (Cochran,
  1965; Rosenbaum, 2005).} We categorized each sample as either a
student convenience sample, a non-student convenience sample, a
probability or representative sample, or another kind of sample
(\(\kappa = .73\)). We categorized the age group(s) in each sample as
children (\(\leq 12\) years), adolescents (13--18 years), or adults
(\(\geq 18\) years; \(\kappa = .64\)).

@@

\textbf{\emph{Predictor Variables.}} For each measure of intergroup
contact, we coded whether it assessed contact with the specific
advantaged outgroup of interest directly---or whether it assessed
contact indirectly by, for example, asking what proportion of someone's
friends were not from the participants' ingroup or what proportion of
residents in someone's neighborhood were from the relevant outgroup
(\(\kappa = .82\)). In addition, we coded whether each predictor
variable assessed direct, qualitatively positive intergroup contact by,
for example, measuring contact quality or cross-group friendship with
the advantaged outgroup(\(\kappa = .82\)).\footnote{We thank John Dixon
  for his suggestion to consider the distinction between contact
  quantity and quality in our analyses. As we added this moderator
  during peer review and as doing so changed the results of our
  moderator analyses, we report our original findings in the SOM.}

@@

We estimated effect sizes with Bayesian random-effects meta-analysis
models, using the \emph{RStan} package (Stan Development Team, 2021) for
the R programming language (R Core Team, 2021), which modeled the
\emph{z}-transformed correlation coefficients with a normal likelihood
function:
\begin{align*} z_{ij} &\sim \text{Normal}(\theta_{ij}, \sigma_{ij}) \\ \theta_{ij} &= \begin{cases} \mu + \beta_j\tau_J & \text{if } I_j = 1 \\ \mu + \beta_j\tau_J + \beta_i\tau_I & \text{if } I_j > 1 \end{cases} \\ \end{align*}
where \(z_{ij}\) is the observed effect size in sample \(i\) of study
\(j\), \(\sigma_{ij}\) is the sample standard error, and \(\theta_{ij}\)
is the estimated effect size. We estimated \(z_{ij}\) as a function of
the mean effect size \(\mu\) and of two varying (random) intercepts,
\(\beta_i\) and \(\beta_j\), with the corresponding standard deviations,
\(\tau_I\) and \(\tau_J\). We used the non-centered parameterization to
model the random effects. For studies that contained only one sample
(\(I_j = 1\)), we only included \(\beta_j\tau_J\), the study-wise
deviation from the mean effect size \(\mu\). For studies that contained
more than one sample (\(I_j > 1\)), we also estimated \(\beta_i\tau_I\),
the sample-wise deviation from the study-specific effect size.

@@

Bayesian inference involves choosing a likelihood function and prior
distributions. A likelihood function links the observed data to one or
more model parameters and states how likely the observed data are given
different values of said model parameters. Prior distributions state how
plausible different values of said model parameters are before
considering the observed data. Bayesian inference applies Bayes' theorem
to update prior distributions in light of the observed data to produce
posterior distributions. In contrast to \emph{p}-values and confidence
intervals, the resulting posterior distributions have a straightforward
interpretation as stating how plausible different values of the model
parameters are given the observed data.\footnote{For an accessible
  introduction, see McElreath (2020).} We report point estimates, based
on the median of posterior samples, and uncertainty intervals, based on
the quantiles of posterior samples, that enclose the 95\% most plausible
estimates. In addition, we report the posterior probability, based on
the proportion of posterior samples below zero, that the mean effect
size is negative.

@@

\hypertarget{deviations-from-the-preregistration}{%
\subsection{Deviations From the
Preregistration}\label{deviations-from-the-preregistration}}

Some aspects of the preregistered protocol proved unrealistic,
impractical, or underspecified. Therefore, we deviated from the
preregistration in the following ways:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  We preregistered that we would update our search every four months
  until we would submit the manuscript. This proved unrealistic as the
  study selection and data collection process for the new studies took
  almost as long. Instead, we concluded our search of electronic
  databases on April 1, 2020---that is, before we started analyzing and
  documenting our findings.
\item
  We preregistered that we would randomly select 100 records to be
  screened by both coders to calculate inter-rater agreement. We did not
  specify, however, what we would do if inter-rater agreement was less
  than acceptable. We decided to refine our coding strategy over three
  samples of 100 records until we achieved acceptable agreement.
\item
  We preregistered that we would use \emph{Google Scholar} to find
  records citing eligible studies. This proved impractical as
  \emph{Google Scholar} does not allow the electronic export of citing
  records. Instead, we used the \emph{Scopus} citation database.
\item
  We preregistered that we would attempt to contact the authors of
  \emph{all} papers with missing effect sizes. Instead, we decided to
  not contact authors for studies published before 2000 as we considered
  it unlikely that the authors still had access to the data.
\item
  We preregistered that we would use a Bayesian two-level random-effects
  meta-analysis model for all preregistered analyses. For the
  association between policy support and collective action, however, we
  had only few studies and found that the two-level model did not result
  in a reliable posterior distribution. Instead, we used a Bayesian
  one-level random-effects model to estimate that association.
\end{enumerate}

@@

As preregistered, we conducted two kinds of robustness checks. First, we
assessed to what extent our findings were sensitive to choosing
narrower, \(\mu \sim \text{Normal}(0, 0.1)\), or wider,
\(\mu \sim \text{Normal}(0, 1)\), prior distributions. Choosing narrower
or wider prior distribution did not affect mean effect size estimates
for perceived injustice (\(\Delta r = -.00, [-.05, .04]\) and
\(\Delta r = .00, [-.04, .05]\)), collective action
(\(\Delta r = -.00, [-.10, .10]\) and \(\Delta r = .00, [-.10, .11]\)),
and policy support (\(\Delta r = -.00, [-.10, .09]\) and
\(\Delta r = .00, [-.09, .10]\)). Second, we assessed to what extent our
findings were sensitive to including or excluding influential studies by
repeating the preregistered analyses \(J\) times while leaving out one
of \(J\) studies each time and by calculating the mean absolute
difference (\emph{MAD}) for the estimated mean effect size across
left-out studies. For perceived injustice
(\(\textit{MAD} = .02, [.01, .04]\)), collective action
(\(\textit{MAD} = .04, [.02, .09]\)), and policy support
(\(\textit{MAD} = .03, [.02, .09]\)), the \emph{MAD} was small. Leaving
out the most influential study, for example, did not change estimates of
the mean effect sizes for the three outcomes
(\(\Delta r = .01, [-.04, .05]\); \(\Delta r = .02, [-.09, .11]\);
\(\Delta r = -.02, [-.10, .08]\)). These analyses showed that our
findings were robust to choosing different prior distributions and to
excluding influential studies.

@@

\begin{figure*}
\centering
\caption{Estimated effect size as a function of cultural distance from the United States, with point estimates and uncertainty intervals for each country}
\includegraphics[scale=1]{../figures/figure-5}
\caption*{\textit{Note.} We had data from too few countries to reach firm conclusions about the direction of these associations for perceived injustice ($\beta = -0.39, [-1.47, 0.77]$; $\Pr (\beta < 0) = 75.4\%$), collective action ($\beta = -0.08, [-1.53, 1.40]$; $\Pr (\beta < 0) = 54.8\%$), and policy support ($\beta = -0.89, [-2.10, 0.67]$; $\Pr (\beta < 0) = 88.9\%$).}
\label{fig:f5}
\end{figure*}

@@

First, we used meta-regression models to examine categorical moderator
variables which we expected to explain heterogeneity in effect sizes
across samples. As Figure 4 shows, we found that the setting of the
study (\(R^2 = 21\%\)), participants' age group (\(R^2 = 12\%\)), and
whether intergroup contact was measured directly or indirectly
(\(R^2 = 10\%\)) explained the most variance across effect sizes.

@@

Third, we used meta-regression trees to discover interactions between
moderator variables that best explained heterogeneity in effect sizes
(Li et al., 2017, 2020). Figure 6 shows the resulting meta-regression
model, which explained more variance across samples than any individual
moderator (\(R^2 = 37\%\)). We found that intergroup contact was
associated with less perceived injustice only in studies that measured
direct, qualitatively positive contact among adults
(\(r = -.12, [-.16, -.09]\)) or in the few studies that did not measure
direct, qualitatively positive contact but in which the groups' relative
inequality stemmed from short-term migration
(\(r = -.23, [-.31, -.16]\)).

@@

Developing methods to detect and correct for publication bias and other
meta-biases is an active area of research, with no single method
outperforming all others (Carter et al., 2019). Following Carter et
al.'s recommendations, we compared results of several methods to adjust
meta-analytic estimates for publication bias: the PET-PEESE estimator
(Stanley \& Doucouliagos, 2014) adjusts meta-analytic estimates by
correcting for the greater risk of smaller studies to not be published;
the \emph{p}-uniform* estimator (van Aert \& Assen, 2018) adjusts
meta-analytic estimates by correcting for the greater risk of
non-significant effects to not be published; and the three-parameter
selection model (3PSM, Vevea \& Hedges, 1995) adjusts meta-analytic
estimates for publication bias by estimating both the probability of a
non-significant effect being published as well as the true average
effect and between-study heterogeneity underlying the observed effect
sizes.\footnote{For an accessible introduction to meta bias correction,
  see Harrer et al. (2021).}

In addition, we ran a subgroup analysis that included only studies that
were not published or not conducted with the intention to examine the
effects of intergroup contact on any of the outcomes---and that,
therefore, can be assumed to have not been affected by publication bias.
As the methods we used do not use Bayesian statistics, we compared
adjusted estimates to unadjusted estimates from a random-effects
meta-analysis estimated with restricted maximum likelihood estimation
using the \emph{metafor} R package (Viechtbauer, 2010).

@@

\hypertarget{additional-analyses}{%
\subsection{Additional Analyses}\label{additional-analyses}}

\begin{figure*}[t!]
\centering
\caption{Posterior distributions for studies that measured direct, qualitatively positive contact among adults}
\includegraphics[scale=1]{../figures/figure-8}
\caption*{\textit{Note.} Black outlines show the posterior distributions from the original, preregistered analyses. \textbf{A} Posterior distributions for the mean correlation coefficients, highlighting the proportion of posterior samples for which $r_\text{mean} < 0$. \textbf{B} Posterior predictive distributions of study-wise correlation coefficients, highlighting the 80\% most common effect sizes, with point estimates for the correlation coefficients for all studies in the sample.}
\label{fig:f8}
\end{figure*}

Our moderator analyses suggested that, for perceived injustice, the
evidence for the `ironic' effect of intergroup contact was stronger in
studies that measured direct, qualitatively positive contact among
adults than in most other studies. Operationalizing intergroup contact
as direct, qualitatively positive contact among adults is closer to the
original argument for the `ironic' effects of intergroup contact and,
therefore, tests a stronger case for the hypothesized relationships. To
test this stronger case, we repeated our original, preregistered
analyses for studies that measured direct, qualitatively positive
contact among adults.
